\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-26 paper.qmd
Updated Figure 2 (need to work on the title)

@@

@@

>>>>

umap_data_two_curvy_with_hb_id <- hb_obj_two_curvy$data_hb_id
df_all_two_curvy <- dplyr::bind_cols(training_data_two_curvy, umap_data_two_curvy_with_hb_id)
df_b_two_curvy <- avg_highd_data(data = df_all_two_curvy, col_start = "x")

@@

distance_df_small_edges_two_curvy <- distance_two_curvy |>
    filter(distance < benchmark_two_curvy)

@@

```{r}
#| label: two-curvy-true-proj

true_model_df <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model.rds")
wireframe_true_model <- read_rds("data/two_curvy_clust/two_curvy_clust_true_model_connections.rds")

data_two_curvy <- training_data_two_curvy |>
  mutate(type = "data")

true_model_two_curvy <- true_model_df |> 
    select(-ID) |> 
  mutate(type = "true model")

# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)

scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
  as_tibble()

scaled_two_curvy <- scale_data_manual(df_model_data_two_curvy, "type") |>
  as_tibble()

scaled_two_curvy_data <- scaled_two_curvy |>
  filter(type == "data") |>
  select(-type)

scaled_two_curvy_data_model <- scaled_two_curvy |>
  filter(type == "true model") |>
  select(-type)

## First projection
projection <- cbind(
    c(-0.08707,-0.06493,0.07017,-0.07349,-0.02955,-0.05286,-0.02765),
    c(0.03398,0.03735,0.09272,0.01871,-0.09224,0.00451,0.08079))

projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled

projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled

model_df <- dplyr::left_join(
  wireframe_true_model, 
  projected_model_df, 
  by = c("from" = "ID"))

axes_obj <- gen_axes(  
  proj = projection * 5,
  limits = 0.15,
  axis_pos_x = -0.1,
  axis_pos_y = -0.075,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0.01)

two_curvy_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  interior_annotation("a") +
  theme(
    legend.position = "none"
  )

## Second projection
projection <- cbind(
    c(-0.02704,0.08281,-0.07671,-0.02996,0.08674,0.06497,-0.02090),
    c(0.00095,0.08031,0.05721,0.10299,0.04816,-0.05837,-0.02226))

projected <- as.matrix(scaled_two_curvy_data) %*% projection_scaled

projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled

model_df <- dplyr::left_join(
  wireframe_true_model, 
  projected_model_df, 
  by = c("from" = "ID"))

axes_obj <- gen_axes(  
  proj = projection * 5,
  limits = 0.15,
  axis_pos_x = -0.1,
  axis_pos_y = -0.075,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0.01)

two_curvy_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  interior_annotation("b") +
  theme(
    legend.position = "none"
  )
```

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

two_curvy_proj_tsne_model1 + two_curvy_proj_tsne_model2 + 
  plot_layout(ncol = 2)
```

### 2024-09-26 paper.tex
Updated Figure 2 (need to work on the title)

>>>>

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two-curvy-true-proj-1.pdf}

\caption{\label{fig-two-curvy-true-proj}Two views of the true model
(blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
S-curve data (black points). The data is spread along the S shape, and
does not vary much from this curve. (The \textbf{langevitour} software
is used to view the data with a tour, and the full video is available at
\url{https://youtu.be/I5GL23vLiw0}).}
